---
title: "HW 1"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Benjamin J. Fox"
  email: "foxbe@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document
---

18/20. Good job Ben

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 21 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

## Preliminary Task
Go to [datacamp.com](https://datacamp.com) and complete the tutorial "Introduction to R".


## Task 1
a) Store the data below in a data frame object named `my_df`.
```{r}
#Creating the vectors for people and preference
people <- c("Alice", "Bob", "Charlie", "Denise", "Erin", "Frank", "Gisele", "Harry")
preference <- c("Twitter", "Facebook", "Twitter", "Instagram", "Facebook", "Twitter", "Twitter", "Instagram")

#Storing the data in a data frame object
my_df <- data.frame(people, preference)

my_df
```

|Person | Preference|
|:------|:----------|
|Alice | Twitter |
| Bob | Facebook |
| Charlie | Twitter |
| Denise | Instagram |
| Erin | Facebook |
| Frank | Twitter |
| Gisele | Twitter |
| Harry | Instagram |

b) What are the dimensions of `my_df`?
The dimensions of "my_df" are people and preference since each entry in one corresponded to a result in the other.

c) Write R code that counts the total number of users.
```{r}
#Get the amount of people and put it in a variable
total_users <- length(people)


#Print it
total_users
```
d) Write R code that counts the total number of users who prefer Twitter.
```{r}
#Use subset to get specifically the people who prefer twitter
twitter = subset(my_df, subset = preference == "Twitter")

#Print that result
twitter
```
e) Write R code that returns what Gisele prefers.
```{r}
#Find where that information is stored
gisele = my_df[7,2]

#Print that data
gisele
```

## Task 2
a) Create a vector named `my_vec` that includes all integers from 1 to 10 (inclusive).
```{r}
#Create that vector
my_vec <- c(1:10)

#Print that vector to test that it worked
my_vec
```
b) Replace the second element in `my_vec` with the number `99`.
```{r}
#Assign 99 to the second value in my_vec
my_vec[2] <- 99

#Print my_vec to make sure the change took
my_vec
```

c) Write a function called `my_func` that prints `This element is greater than 5!` if an element in `myvec` is greater than 5, and `This element is NOT greater than 5!` otherwise.
```{r}
#Make a function
my_function <- function(vector){
  #For-loop to sort through the vector
  for (i in my_vec) {
    #If-else ladder to print the correct statement
    if (i > 5) {
      print("This element is greater than 5!")
    } else {
      print("This element is NOT greater than 5!")
    }
  }
}

#Test function
my_function(my_vec)
```

## Task 3
a. Write a function named `get_fibonacci` that returns the n'th value in the [fibonacci sequence](https://www.mathsisfun.com/numbers/fibonacci-sequence.html). For example, `get_fibonacci(5)` should return 3; `get_fibonacci(7)` should return 8 and so on. ***Hint:** It is possible to call the same function from inside it. E.g., you can call `get_fibonacci()` from inside `get_fibonacci()`. Like inception!*
```{r}
#Create the function
get_fibonacci <- function(x){
  #Assign the first two values of the sequence
  p <- 0
  q <- 1
  
  #For loop to find the rest of the fibonacci sequence
  for (i in 1:x){
    #Calculate the next number in the sequence
    next_num <- p + q
    
    #Reassign numbers accordingly
    p <- q
    q <- next_num
  }
  
  return(p)
}
```

b. Using your function in part (a), find and return the 19th value in the fibonacci sequence.
```{r}
#Find  the wanted number
get_fibonacci(5)
# 0, 1, 1, 2, 3, 5, 8, 13, 21, 34, 55, 89, 144. This is the fibonacci sequence. get_fibonacci(5)should return 3. 

```

<br>
You are done! 🏁 Don't forget to `commit` and `push` your .Rmd file to your Github clone repository before the due date.
<br>

