---
title: "HW 2: Visualization, Base R"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Benjamin J. Fox"
  email: "foxbe@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---
17/20. Good job. Sound analysis.

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 28 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality). 

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,
```{r}
head(Cars93)
```

# Base R
## Task 1
a) Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.
```{r}
#Creating a frequency table
type_frequency <- table(Cars93$Type)

#Finding the length of that table to make relative frequency easier
type_length <- length(Cars93$Type)

#Finding the relative frequency table
type_percent_frequency <- round(type_frequency/type_length, 3)
```
b) Create a frequency bar graph for the variable `Type`. You can have count on the y axis.
```{r}
barplot(type_frequency, 
        col = c("red", "orange", "yellow", "green", "blue", "purple"), 
        main = "Type Frequency Bar Graph", cex.axis = 1.2, 
        cex.names = 1.2, cex.main = 1.2)
```

*	Your axes should be labeled and legible
*	The bars should be of different colors
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

c) Create a pie chart of the variable `Type`.
```{r}
pie(type_frequency, main = "Pie Chart of Type Frequency", 
    cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2, 
    col = c("red", "orange", "yellow", "green", "blue", "purple"))
```

* The plot should include a title using the main argument.
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

d) What is the mode of the variable `Type`? Comment.
What is the mode? (-2)

## Task 2 
a) Create a histogram of the variable `Horsepower`.
```{r}
#Create horsepower histogram
hist(Cars93$Horsepower, col = "orchid", xlab = "Horsepower", 
          cex.lab = 1.2, cex.axis = 1.2,  cex.main = 1.2)
```

*	Your axes should be labeled
*	The hist body should have a color (your choice, not too dark to obscure lines) 
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes should be set to 1.2

b) Create a second histogram with breaks at 25 HP increments instead of 50.
```{r}
#Create basically the same histogram and tweak the bin width
hist(Cars93$Horsepower, col = "orchid", xlab = "Horsepower", 
          cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2, breaks = 10)
```

* Use the breaks argument to define the classes
* Other setting should be as in part (a)

c) Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.
```{r}
#Find the length of the horsepower data
horsepower_length <- length(Cars93$Horsepower)

#Find the square root of the length
horsepower_ideal_classes <- round(sqrt(horsepower_length))

horsepower_ideal_classes

#Since horsepower ideal classes = 10, and when binwidth = 25 you have 11 bins, when binwidth = 50 you have 6 bins. Therefore the second histogram most closely follows the correct rule of thumb.
```


## Task 3 
a) What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?
```{r}
#Print Passengers
Cars93$Passengers

#It is a numerical discrete variable. It's a bad idea to make this variable into a histogram because you won't need to sort them into ranges both because they're integers and the width of the possible values isn't large enough to warrant using a histogram.
```

b) Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with no observations on the plot.)
```{r}
#I'll make a table to sort things out a bit easier when I get to making the bar plot
passenger_frequency <- table(Cars93$Passengers)

#A bar graph is better, so I'll make that.
barplot(passenger_frequency, main = "Frequency of Passengers", 
        xlab = "Amount of Passengers", ylab = "Frequency", 
        col = c("pink", "#ff8888", "#ffffaa", "#aaffaa", "#aaccff", "#aa66aa"), 
        cex.axis = 1.2, cex.lab = 1.2, cex.main = 1.2)

# This bar plot should show the frequency of cars that carry 1 and 3 passengers as 0 (-2)
```

## Task 4 
a) Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.
```{r}
hist(Cars93$Length, col = "orchid", xlab = "Length", 
          cex.lab = 1.2, cex.axis = 1.2,  cex.main = 1.2)
```

b) What is the shape of the distribution?
It is a largely normally distributed function, with a very slight skew right. 

c) Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)?
```{r}
#What is the mean?
mean(Cars93$Length)
#I think that 183.2043 makes sense, this is around the mid point of the histogram and is also where the graph has the highest frequency

#What is the median?
median(Cars93$Length)
#The median is just 183, which also makes sense, that's around the middle, and the mode, plus it's only a little smaller than what the mean is so it helps the belief that the data is normally distributed.
```

You are done! ðŸ Don't forget to commit and push your .Rmd file to your Github repository before the due date.


