---
title: "HW 3: Visualization, Ggplot2 & Maps"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Benjamin J. Fox"
  email: "foxbe@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** October 12 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: This time we will practice visualization using ggplot2, and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(ggplot2) # keep in mind you have to install ggplot2 first via install.packages() if it's not installed in your machine
library(MASS)
data(Cars93)
head(Cars93)
```

# Ggplot2

## Task 1
Create a simple scatterplot using ggplot2 where x axis shows `EngineSize`, y axis shows `MPG.city`.

* Use blue color for your dots.
* Use size = 2 for the dot sizes.
* Use xlim and ylim to make sure your x and y axis ranges make sense.
* Add a title, subtitle, x axis label, y axis label, and a caption using title, subtitle, caption options inside `+ labs()`.

 
```{r}
#Mapped out the requested with unhelpful text cause I wasn't sure what to write where.
ggplot(Cars93, aes(EngineSize, MPG.city)) + geom_point(size = 2, colour = "blue") + labs(title = "Engine Size to Miles Per Gallon in Cities", subtitle = "This is the chart you asked me to make", caption = "This is how we do it") + xlab("Engine Size") + ylab("Miles Per Gallon (City)")
```

## Task 2
Recreate the scatterplot in Task 1. Add the following functionality:

* Use a different color for each car `Type`.
```{r}
#Copy-pasted the graph and colored the necessary factor variable.
ggplot(Cars93, aes(EngineSize, MPG.city, colour = Type)) + geom_point(size = 2) + labs(title = "Engine Size to Miles Per Gallon in Cities", subtitle = "This is the chart you asked me to make", caption = "This is how we do it") + xlab("Engine Size") + ylab("Miles Per Gallon (City)")
```

## Task 3
Recreate the scatterplot in Task 2. Add the following functionality:

* Fit and overlay a smooth curve.
* Make the curve line red, and dashed. (Hint linetype is the option you can use to set line style as 'dashed')

```{r}
#Added a smooth dashed line over the previous data, and moved the color-type thing so only the dots would be colored
ggplot(Cars93, aes(EngineSize, MPG.city)) + geom_point(size = 2, aes(colour = Type)) + labs(title = "Engine Size to Miles Per Gallon in Cities", subtitle = "This is the chart you asked me to make", caption = "This is how we do it") + xlab("Engine Size") + ylab("Miles Per Gallon (City)") + geom_smooth(method = "loess", se = F, linetype = "dashed", color = "red")
```

## Task 4
Create a barplot showing count of cars by `AirBags` using ggplot2.

* `geom_bar()` is the function that creates a barplot.
* Use a different color for each bar.
* Add a title, subtitle, x axis label, y axis label, and a caption using title, subtitle, caption options inside `+ labs()`.

```{r}
#Made a bar graph and colored them so it's Halloween themed.
ggplot(Cars93, aes(AirBags)) + geom_bar(fill = c("#ff7700", "#00aa00", "#aa0088")) + labs(title = "Frequency of Airbag Implementation", subtitle = "Data gathered from Cars93 dataset", caption = "Sponsored by Subway lmao") + xlab("Airbag Implementation") + ylab("Frequency")
```

## Task 5
Recreate the barplot in Task 4 for different car `Origin`, such that you create one barplot for each `Origin` category and display all on the same graph.

* You can have `Origin` as the "x" variable by setting `x=Origin.`
* Use `fill=AirBags` to get a different color (and bar) by `AirBags`.
* in `geom_bar(), use `position='dodge'` to have the barplots side by side.

```{r}
#Did the same bar graph except I added another variable and made sure it was graphed in the desired format.
ggplot(Cars93, aes(x=Origin, fill = AirBags)) + geom_bar(position = "dodge") + labs(title = "Frequency of Airbag Implementation", subtitle = "Data gathered from Cars93 dataset", caption = "Sponsored by Subway lmao") + xlab("Airbag Implementation Between Regions") + ylab("Frequency")
```

# Maps
Load Australia's state map data via running:
```{r}
# install the packages below by uncommenting the lines. You need to do this only once.
#install.packages('ozmaps') # downloads the map needed
#install.packages('sf') # package used to analyze spatial data
library(ozmaps)
library(sf)
au_states <- ozmaps::ozmap_states # read in spatial dataframe object

# Display the data
head(au_states)
au_states
```
Notice that `NAME` includes the name of a state, `geometry` includes spatial coordinate data for each state polygon (area).

## Task 1
Create a map of Australia using `ggplot2`.
```{r}
#Just used the basic requested command
ggplot(au_states) + geom_sf()
```

## Task 2
Modify your map such that each state is displayed with a different color. (Hint: use the fill option as an aes setting with variable `NAME`.)
```{r}
#Copypasted the command and added an aes variable to note the different Australian states
ggplot(au_states, aes(fill = NAME)) + geom_sf()
```
## Task 3
Read in the capital city dataframe showing each capital and its latitude and longitude information by running:
```{r}
au_capitals <- tibble::tribble( 
  ~city,           ~lat,     ~lon,
  "Sydney",    -33.8688, 151.2093,  
  "Melbourne", -37.8136, 144.9631, 
  "Brisbane",  -27.4698, 153.0251, 
  "Adelaide",  -34.9285, 138.6007, 
  "Perth",     -31.9505, 115.8605, 
  "Hobart",    -42.8821, 147.3272, 
  "Canberra",  -35.2809, 149.1300, 
  "Darwin",    -12.4634, 130.8456, 
)
```

Display the capitals as points on your map from Task 1. (Hint: combine `geom_sf()` with `+ geom_point()`)

* Use `col = "red"` such that points are displayed in red color.
* Use `size=2` to make points larger.

```{r}
#Did the same command and added dots for the state capitals. I also put the data source in each geom command in order to make sure they wouldn't interfere with each other.
ggplot() + geom_sf(data = au_states, aes(fill = NAME)) + geom_point(data = au_capitals, aes(lon, lat), col = "red", size = 2)
```


## Task 4
Let's take a look at a different projection output. Change your map's projection to epsg 3112 by adding `+ coord_sf()`. Recall the option that goes inside `coord_sf()` is `st_crs(4_digit_code)`
```{r}
#I honestly don't fully don't know what you want here, this just seems to be giving me the map from before and putting coord_sf after geom_sf is just throwing errors
default_crs <- sf::st_crs(4326)
ggplot() + coord_sf(default_crs) + geom_sf(data = au_states, aes(fill = NAME)) + geom_point(data = au_capitals, aes(lon, lat), col = "red", size = 2)
```

## Task 5
Lastly, let's display the same map using `mapview()`. Load the `mapview` package.

```{r}
#install.packages('mapview') # interactive display
library(mapview)
```
Create an interactive map of Australia using `mapview()`

```{r}
#Just put in the command
mapview(au_states)
```

## Task 6
Let's display the capitals on the map like before. First convert the `au_capitals` dataframe to spatial dataframe object by running:
```{r}
capitals_geo <- st_as_sf(au_capitals, coords = c("lon", "lat"))
```

Now add the new spatial capitals object via + (that is, you can use the following syntax to add a layer: `mapview() + mapview()`)

* You can use the option `col.regions='red'` to make capitals appear in color red, and `legend=F` in second `mapview()` to remove the uninformative second set of legend. 

```{r}
#Just tacked the other command on and added the requested options.
mapview(au_states) + mapview(capitals_geo, col.regions="red", legend=F)
```



You are done! ðŸ Don't forget to commit and push your .Rmd file to your Github repository before the due date.


